Notes on Design and Implementation of Anglican

### 2015-07-28 paper/paper.tex
typo
author

\title{Notes on Design and Implementation of Anglican}

@@

\begin{abstract}
	In these notes, are explain the rationale behind the
	macro-compiled implementation of Anglican, outline
	design choices, and describe important implementation
	aspects.
\end{abstract}

@@

In implementations of functional programming languages stack
growth is avoided through \textit{tail call optimization} (TCO).
However, Clojure does not support a general form of TCO, and
CPS-transformed code that creates deeply nested calls will
easily exhaust the stack. Anglican employs a workaround called
\textit{trampolining} --- instead of inserting a continuation
call directly, the transformer always wraps the call into a
\textit{thunk}, or parameterless function. The thunk is returned
and called by the trampoline (Clojure provides function
\texttt{trampoline} for this purpose) --- this way the
computation continues but the stack is collapsed on every
continuation call. Function \texttt{continue} implements the
wrapping.

To realize the potential danger of code size explosion, consider
CPS transformation of code
\begin{lstlisting}[frame=single,language=Anglican, basicstyle=\scriptsize\ttfamily, showstringspaces=false]
(if (adult? person)
  (if (male? person)
	 (choose-beer)
	 (choose_wine))
  (choose-juice))
\end{lstlisting}
with continuation 
\begin{lstlisting}[frame=single,language=Anglican, basicstyle=\scriptsize\ttfamily, showstringspaces=false]
(fn [choice _]
  (case (kind choice)
    :beer (beer-jar choice)
    :wine (wine-glass choice)
    :juice (juice-bottle choice)))
\end{lstlisting}
The code of continuation, represented by an \texttt{fn} form, will be
repeated three times. In general, CPS code can grow extremely
large if symbolic continuations are inserted repeatedly.

>>>>

\title{Notes on Design and Implementation of Anglican}
\author{David Tolpin\\dtolpin@robots.ox.ac.uk}

@@

\begin{abstract}
	In these notes, I explain the rationale behind the
	macro-compiled implementation of Anglican, outline
	design choices, and describe important implementation
	aspects.
\end{abstract}

@@

In implementations of functional programming languages stack
growth is avoided through \textit{tail call optimization} (TCO).
However, Clojure does not support a general form of TCO, and
CPS-transformed code that creates deeply nested calls will
easily exhaust the stack. Anglican employs a workaround called
\textit{trampolining} --- instead of inserting a continuation
call directly, the transformer always wraps the call into a
\textit{thunk}, or parameterless function. The thunk is returned
and called by the trampoline (Clojure provides function
\texttt{trampoline} for this purpose) --- this way the
computation continues, but the stack is collapsed on every
continuation call. Function \texttt{continue} implements the
wrapping.

To realize potential danger of code size explosion, consider
CPS transformation of code
\begin{lstlisting}[frame=single,language=Anglican, basicstyle=\scriptsize\ttfamily, showstringspaces=false]
(if (adult? person)
  (if (male? person)
	 (choose-beer)
	 (choose_wine))
  (choose-juice))
\end{lstlisting}
with continuation 
\begin{lstlisting}[frame=single,language=Anglican, basicstyle=\scriptsize\ttfamily, showstringspaces=false]
(fn [choice _]
  (case (kind choice)
    :beer (beer-jar choice)
    :wine (wine-glass choice)
    :juice (juice-bottle choice)))
\end{lstlisting}
The code of continuation, represented by an \texttt{fn} form, will be
repeated three times. In general, CPS code can grow extremely
large if symbolic continuations are inserted repeatedly.
